---
title: "within income decile include vs exclude income<$10000"
author: "Chongjun Liao"
date: "11/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The second analysis is similar to the first one, instead of fitting the regression model on the whole dataset, we fit a regression model within each household income decile and see whether including applicants who reported household income below `$10,000` would change the regression results.
In figure 1, we can see the regression on the lowest decile explains more variation in household income by SAT scores after including applicants. While the results of regressions on other deciles have no significant change.  
In figure 2, we also find the regression on the lowest decile performs better.  
```{r,message=FALSE}
# topics by decile
library(dplyr)
setwd("~/Desktop/DS4SI/final project/replication code & data/data")
merged <- read.csv("merged_final.csv")

income_df <- merged %>%
  filter(FAMILY_INCOME > 10000)
fi_cuts <- quantile(na.omit(income_df$FAMILY_INCOME),
                    probs = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))
income_df$fi <- income_df$FAMILY_INCOME

income_df <- income_df %>%
  mutate(percentile = case_when(fi <= fi_cuts[1] ~ 1,
                                fi > fi_cuts[1] & fi <= fi_cuts[2] ~ 2,
                                fi > fi_cuts[2] & fi <= fi_cuts[3] ~ 3,
                                fi > fi_cuts[3] & fi <= fi_cuts[4] ~ 4,
                                fi > fi_cuts[4] & fi <= fi_cuts[5] ~ 5,
                                fi > fi_cuts[5] & fi <= fi_cuts[6] ~ 6,
                                fi > fi_cuts[6] & fi <= fi_cuts[7] ~ 7,
                                fi > fi_cuts[7] & fi <= fi_cuts[8] ~ 8,
                                fi > fi_cuts[8] & fi <= fi_cuts[9] ~ 9,
                                fi > fi_cuts[9] ~ 10))
just_essays_income <- replicate(10,NA)
#1
sat_perc <-  income_df %>%
  filter(percentile == 1)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[1]=summary(mod)$adj.r.squared
#2
sat_perc <-  income_df %>%
  filter(percentile == 2)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[2]=summary(mod)$adj.r.squared
#3
sat_perc <-  income_df %>%
  filter(percentile == 3)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[3]=summary(mod)$adj.r.squared
#4
sat_perc <-  income_df %>%
  filter(percentile == 4)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[4]=summary(mod)$adj.r.squared
#5
sat_perc <-  income_df %>%
  filter(percentile == 5)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[5]=summary(mod)$adj.r.squared
#6
sat_perc <-  income_df %>%
  filter(percentile == 6)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[6]=summary(mod)$adj.r.squared
#7
sat_perc <-  income_df %>%
  filter(percentile == 7)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[7]=summary(mod)$adj.r.squared
#8
sat_perc <-  income_df %>%
  filter(percentile == 8)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[8]=summary(mod)$adj.r.squared
#9
sat_perc <-  income_df %>%
  filter(percentile == 9)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[9]=summary(mod)$adj.r.squared
#10
sat_perc <-  income_df %>%
  filter(percentile == 10)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income[10]=summary(mod)$adj.r.squared
```

```{r, error = TRUE}
just_essays_dict <- replicate(10,NA)
#1
sat_perc <-  income_df %>%
  filter(percentile == 1)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[1]=summary(mod)$adj.r.squared
#2
sat_perc <-  income_df %>%
  filter(percentile == 2)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[2]=summary(mod)$adj.r.squared
#3
sat_perc <-  income_df %>%
  filter(percentile == 3)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[3]=summary(mod)$adj.r.squared
#4
sat_perc <-  income_df %>%
  filter(percentile == 4)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[4]=summary(mod)$adj.r.squared
#5
sat_perc <-  income_df %>%
  filter(percentile == 5)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[5]=summary(mod)$adj.r.squared
#6
sat_perc <-  income_df %>%
  filter(percentile == 6)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[6]=summary(mod)$adj.r.squared
#7
sat_perc <-  income_df %>%
  filter(percentile == 7)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[7]=summary(mod)$adj.r.squared
#8
sat_perc <-  income_df %>%
  filter(percentile == 8)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[8]=summary(mod)$adj.r.squared
#9
sat_perc <-  income_df %>%
  filter(percentile == 9)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[9]=summary(mod)$adj.r.squared
#10
sat_perc <-  income_df %>%
  filter(percentile == 10)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict[10]=summary(mod)$adj.r.squared
```

```{r}
income_df <- merged
fi_cuts <- quantile(na.omit(income_df$FAMILY_INCOME),
                    probs = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))
income_df$fi <- income_df$FAMILY_INCOME

income_df <- income_df %>%
  mutate(percentile = case_when(fi <= fi_cuts[1] ~ 1,
                                fi > fi_cuts[1] & fi <= fi_cuts[2] ~ 2,
                                fi > fi_cuts[2] & fi <= fi_cuts[3] ~ 3,
                                fi > fi_cuts[3] & fi <= fi_cuts[4] ~ 4,
                                fi > fi_cuts[4] & fi <= fi_cuts[5] ~ 5,
                                fi > fi_cuts[5] & fi <= fi_cuts[6] ~ 6,
                                fi > fi_cuts[6] & fi <= fi_cuts[7] ~ 7,
                                fi > fi_cuts[7] & fi <= fi_cuts[8] ~ 8,
                                fi > fi_cuts[8] & fi <= fi_cuts[9] ~ 9,
                                fi > fi_cuts[9] ~ 10))
just_essays_income1 <- replicate(10,NA)
#1
sat_perc <-  income_df %>%
  filter(percentile == 1)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[1]=summary(mod)$adj.r.squared
#2
sat_perc <-  income_df %>%
  filter(percentile == 2)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[2]=summary(mod)$adj.r.squared
#3
sat_perc <-  income_df %>%
  filter(percentile == 3)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[3]=summary(mod)$adj.r.squared
#4
sat_perc <-  income_df %>%
  filter(percentile == 4)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[4]=summary(mod)$adj.r.squared
#5
sat_perc <-  income_df %>%
  filter(percentile == 5)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[5]=summary(mod)$adj.r.squared
#6
sat_perc <-  income_df %>%
  filter(percentile == 6)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[6]=summary(mod)$adj.r.squared
#7
sat_perc <-  income_df %>%
  filter(percentile == 7)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[7]=summary(mod)$adj.r.squared
#8
sat_perc <-  income_df %>%
  filter(percentile == 8)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[8]=summary(mod)$adj.r.squared
#9
sat_perc <-  income_df %>%
  filter(percentile == 9)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[9]=summary(mod)$adj.r.squared
#10
sat_perc <-  income_df %>%
  filter(percentile == 10)
gamma <- as.matrix(log(sat_perc[,6:75]))
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_income1[10]=summary(mod)$adj.r.squared
```

```{r}
just_essays_dict1 <- replicate(10,NA)
#1
sat_perc <-  income_df %>%
  filter(percentile == 1)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[1]=summary(mod)$adj.r.squared
#2
sat_perc <-  income_df %>%
  filter(percentile == 2)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[2]=summary(mod)$adj.r.squared
#3
sat_perc <-  income_df %>%
  filter(percentile == 3)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[3]=summary(mod)$adj.r.squared
#4
sat_perc <-  income_df %>%
  filter(percentile == 4)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[4]=summary(mod)$adj.r.squared
#5
sat_perc <-  income_df %>%
  filter(percentile == 5)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[5]=summary(mod)$adj.r.squared
#6
sat_perc <-  income_df %>%
  filter(percentile == 6)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[6]=summary(mod)$adj.r.squared
#7
sat_perc <-  income_df %>%
  filter(percentile == 7)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[7]=summary(mod)$adj.r.squared
#8
sat_perc <-  income_df %>%
  filter(percentile == 8)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[8]=summary(mod)$adj.r.squared
#9
sat_perc <-  income_df %>%
  filter(percentile == 9)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[9]=summary(mod)$adj.r.squared
#10
sat_perc <-  income_df %>%
  filter(percentile == 10)
gamma <- as.matrix(sat_perc[,76:167])
mod <- lm(RSAT_TOTAL_SCORE ~ gamma,data = sat_perc)
just_essays_dict1[10]=summary(mod)$adj.r.squared
```



```{r, fig.cap="topics on income"}
par(mfrow=c(1,2))
barplot(just_essays_income)
title("exclude")
barplot(just_essays_income1)
title("included")
```  



```{r,fig.cap="dictionary patterns on income"}
par(mfrow=c(1,2))
barplot(just_essays_dict)
title("excluded")
barplot(just_essays_dict1)
title("included")
```
